
---
title: "Differential methylation report"
author: "iCONICS"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
    html_document:
        theme: lumen
        toc: false
        toc_float:
            collapsed: false
        df_print: paged
params:
    configfile: null
---

<style>
    body .main-container {
        max-width: 100%;
    }
    h1.title {
        text-align: center;
        color: DarkRed;
    }
    h1, h2, h3 {
        color: DarkBlue;
    }
    hr {
        height: 10px;
        border: 0;
        box-shadow: 0 10px 10px -10px #8c8b8b inset;
    }
    table {
        width: 100%;
    }
    th {
        background-color: LightBlue;
    }
    tr:nth-child(odd) {
        background-color: #f2f2f2;
    }
    tr:hover {
        background-color: LightGreen;
    }
</style>


```{r setup, include=F}
knitr::opts_chunk$set(echo=T)
library(ggplot2)
library(DT)
config <- yaml::yaml.load_file(params$configfile)
```

<hr>

## Parameters


```{r summary, echo=F, results="asis"}
samples <- names(config$cpg_report_of_sample)

df_covariates <- read.table(config$covariates_tsv, sep="\t", header=T)

df_params <- data.frame(
    "Number of samples"=length(samples),
    "Reference build"=config$ref_build,
    "Merge complementary CpGs"=config$merge_complementary_cpgs,
    "Minimum number of measured samples otherwise CpG is removed"=config$min_samples,
    "dmrseq 'minNumRegion' argument: minimum number of CpGs in a DMR"=config$min_cpgs,
    "dmrseq 'maxGap' argument: maximum basepair distance between neighboring CpGs to be included in the same DMR"=config$max_gap,
    "dmrseq 'cutoff' argument"=config$cutoff,
    "dmrseq 'smooth' argument"=config$smooth,
    "dmrseq 'bpSpan' argument: smoothing window (bp)"=config$bp_span,
    "dmrseq 'minInSpan' argument: minimum number of CpGs in a smoothing window"=config$min_in_span,
    "dmrseq 'maxGapSmooth' argument: maximum basepair distance between neighboring CpGs when smoothing"=config$max_gap_smooth,
    "dmrseq 'maxPerms' argument: maximum number of permutations"=config$max_perms,
    "Test covariate"=config$test_covariate,
    "Covariates to be adjusted for"=if(length(config$adjust_covariates) == 0) "NA" else paste(config$adjust_covariates, collapse=", "),
    "dmrseq 'matchCovariate' argument: a covariate to be blocked when permutating, only permutations with balanced composition of this covariate will be used (2-group comparison only)"=if (is.null(config$match_covariate)) "NA" else config$match_covariate,
    check.names=F)
knitr::kable(t(df_params), align="c", col.names=NA)
```

<hr>

## Samples
```{r samples, echo=F, comment=""}
writeLines(samples)
```

<hr>

## Covariates

```{r covariates, echo=F, comment=""}
knitr::kable(df_covariates, align="c")
```

<hr>

## PCA

```{r, echo=F}
pca_svg <- file.path(config$outdir, "pca_of_methylation_ratios.svg")
knitr::include_graphics(pca_svg)
```

<hr>

## Hierarchical clustering

Using euclidean distance to compute the dissimilarity matrix and the Ward2 criterion for the clustering.

```{r, echo=F}
dendrogram_svg <- file.path(config$outdir, "dendrogram_of_methylation_ratios.svg")
knitr::include_graphics(dendrogram_svg)
```

<hr>

## Annotated DMRs which qval <= 0.05

```{r, echo=F, warning=F}
annotated_dmrs_bed <- file.path(config$outdir, "annotated_dmrs.bed")
# Load DMRs
# quote="" is required because of the "3' end" string in the table
df_annotated_dmrs <- read.table(annotated_dmrs_bed, sep="\t", header=T, quote="")

# Keep only DMRs which qval <= 0.05
df_annotated_dmrs <- df_annotated_dmrs[df_annotated_dmrs$qval <= 0.05, ]

# Create interactive table
DT::datatable(df_annotated_dmrs, options=list(autoWidth=T, pageLength=nrow(df_annotated_dmrs)))
```
